---
title: "02 Presenting the data set"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# some file setup parameters
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999) # display numbers in non-scientific format
```


# The Loans Data Set
- ------------------------------------------------------------------------------

The 'Lending Club' is a financial service company headquarted in San Francisco. Until 2019, their business model was peer-to-peer lending: They connected people who wanted to lend money to people who wanted to borrow money. The following data set contains contains real data from past loans!

Links:  
https://www.lendingclub.com/  
https://en.wikipedia.org/wiki/LendingClub  
Disgruntled creditor/investor:
https://www.youtube.com/watch?v=bSeOrUIfLSc

```{r}
data_url <- "https://raw.githubusercontent.com/staehlo/mlr3_and_tidymodels/main/example_data_set.csv"
writeLines(readLines(data_url, 4))
```

| Header         | Explanation
| -------------- | ------------
| int_rate       | interest rate in percent
| loan_amnt      | requested loan amount in US Dollar
| term           | number of month till total pay back [1]
| grade          | loan security rating computed the Lending Club [2]
| home_ownership | housing situation of the borrower [3]
| annual_inc     | annual gross income of the borrower in US Dollar
| purpose        | intended use of money as indicated by the borrower

Details:  
[1] term: Borrowers have to make one back payment per month throughout the term.
[2] grade: The rating is calculated by Lending Club's own algorithm. The main input parameter is the borrower's FICO score (comparable to Germany's "Schufa-Abfrage"). Ratings go from "A" (pay back on time is very likely) till "G" (some risk that the borrower won't be able to pay the money back).
[3] home_ownership: Categories are mortgage, none, other, own, rent

Possible analyses:

(1) Regression: Train a model to predict Lending Club's interest rates
                (based on loan amount, home ownership, annual income, and purpose)

(2) Classification: Term has only two values: 36 and 60 month (3 or 5 years)
                    Train a model to predict Lending Club's proposed payback times.
