---
title: "02 The example data set"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# some file setup parameters
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999) # display numbers in non-scientific format
```


# The Loans Data Set
- ------------------------------------------------------------------------------

The 'Lending Club' is a financial service company headquarted in San Francisco. Until 2019, their business model was peer-to-peer lending: They connected people who wanted to lend money to people who wanted to borrow money. The following data set contains contains real data from past loans!

Links:  
https://www.lendingclub.com/  
https://en.wikipedia.org/wiki/LendingClub  
Disgruntled creditor/investor:
https://www.youtube.com/watch?v=bSeOrUIfLSc

```{r}
data_url <- "https://raw.githubusercontent.com/staehlo/mlr3_and_tidymodels/main/example_data_set.csv"
writeLines(readLines(data_url, 4))
```

| Header         | Explanation
| -------------- | ------------
| int_rate       | interest rate in percent
| loan_amnt      | requested loan amount in US Dollar
| term           | number of month till total pay back [1]
| grade          | loan security rating computed the Lending Club [2]
| home_ownership | housing situation of the borrower [3]
| annual_inc     | annual gross income of the borrower in US Dollar
| purpose        | intended use of money as indicated by the borrower

Details:  
[1] term: Borrowers have to make one back payment per month throughout the term.
[2] grade: The rating is calculated by Lending Club's own algorithm. The main input parameter is the borrower's FICO score (comparable to Germany's "Schufa-Abfrage"). Ratings go from "A" (pay back on time is very likely) till "G" (some risk that the borrower won't be able to pay the money back).
[3] home_ownership: Categories are mortgage, none, other, own, rent

Possible analyses:

(1) Regression: Train a model to predict Lending Club's interest rates
                (based on loan amount, home ownership, annual income, and purpose)

(2) Classification: Term has only two values: 36 and 60 month (3 or 5 years)
                    Train a model to predict Lending Club's proposed payback times.


Load data
```{r}
loans <- read.csv2(file = data_url, header = TRUE, sep = ",", dec=".")
rm(data_url)
str(loans)
```

Check the income distribution
```{r}
plot(loans$annual_inc, col = "maroon",
     xlab = "random order", ylab = "annual income in USD")
hist(loans$annual_inc, 50, col = "maroon")
```

We will focus on people with an annual income below 200'000 USD
```{r}
loans <- loans[loans$annual_inc < 200000,]
plot(loans$annual_inc, col = "maroon",
     xlab = "random order", ylab = "annual income in USD")
hist(loans$annual_inc, 50, col = "maroon", xlab = "annual income in USD")
```

Check the frequency of the home owner categories
```{r}
table(loans$home_ownership)
```

We will ignore the three cases with no indicated home ownership
```{r}
loans <- loans[loans$home_ownership != "NONE",]
loans <- droplevels(loans) # remove the no longer used "NONE" level
table(loans$home_ownership)
```

# Preprocessing
```{r}
loans$home_ownership <- as.factor(loans$home_ownership)
loans$purpose <- as.factor(loans$purpose)
loans$term <- as.factor(loans$term) # for the classification task

# convert grade to numeric:
loans$numeric_grade <- match(loans$grade, LETTERS)

# loans$grade <- as.factor(loans$grade)
# loans$monthly_payback <-round(loans$loan_amnt / loans$term, 2)

str(loans)
```

# Exploratory Plotting
```{r}
set.seed(123)
tmp <- loans[sample(nrow(loans), 1000),]

plot(tmp$annual_inc, tmp$int_rate, col = "maroon",
     xlab = "annual income in USD", ylab = "interest rate in percent")

plot(tmp$loan_amnt, tmp$int_rate, col = "maroon",
     xlab = "requested loan in USD", ylab = "interest rate in percent")

boxplot(int_rate ~ home_ownership, tmp, col = "maroon",
        xlab = "", ylab = "interest rate in percent")

par.default <- par()
par(mar = c(7, 4.1, 4.1, 2.1))
x <- boxplot(int_rate ~ purpose, tmp, col = "maroon",
        xlab = "", ylab = "interest rate in percent", xaxt = "n")
nlevels = length(levels(tmp$purpose))
text(x = (1:nlevels)-0.4, y = 2, labels = x$names, xpd=NA, srt = 45)
par(par.default)

x <- boxplot(int_rate ~ grade, tmp, col = "maroon",
             ylab = "interest rate in percent")

rm(x, nlevels, tmp, par.default)
```