---
title: "03 Linear regression"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# some file setup parameters
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999) # display numbers in non-scientific format
```

(1) Regression: Train a model to predict Lending Club's interest rates
                (based on loan amount, home ownership, annual income, and purpose)

Dependent variable:

* int_rate

Predictor variables:

* loan_amnt (integer)
* annual_inc (numeric)
* numeric_grade (integer)
* home_ownership (factor)
* purpose (factor)


********************************************************************************
# Standalone: Using lm from the stats package
********************************************************************************

## Define train and test rows
```{r}
set.seed(123)
train <- sample(1:nrow(loans), 0.8 * nrow(loans))
test <- setdiff(1:nrow(loans), train)
```

## Fit linear model
```{r}
lm.fit <- lm(int_rate ~ loan_amnt + annual_inc + numeric_grade
             + home_ownership + purpose, data = loans, subset = train)
lm.fit
summary(lm.fit) # display maybe a bit confusing

# for better table overview of results
library(car)
Anova(lm.fit, type = "III")
```

## Evaluation
```{r}
lm.predict <- predict.lm(lm.fit, newdata = loans[test,])

# Variance explained in training data
summary(lm.fit)$r.squared

# Variance explained in test data  (not sure if this step is correct)
(cor(loans[test,]$int_rate, lm.predict))^2

# Root mean squared error (RMSE) of training data set:
mean(sqrt((loans[train,]$int_rate - lm.fit$fitted.values)^2))
# Root mean squared error (RMSE) of test data set:
mean(sqrt((loans[test,]$int_rate - lm.predict)^2))

# plot true versus true
plot(lm.predict, loans[test,]$int_rate, col = "maroon",
     main = "Test data set",
     xlab = "predicted interest rate", ylab = "true interest rate")
abline(0, 1, col = "blue")
```
> r.squared (train) 90.234
> root mean squared error (train) 0.9151408
> r.squared (test)  89.829
> root mean squared error (test) 0.9151887

```{r}
# little clean-up
rm(test, train)
```


********************************************************************************
# mlr3
********************************************************************************

## Load packages
```{r}
# install.packages("mlr3verse")
library("mlr3")
```

## Construct task
```{r}
mlr3task = as_task_regr(loans, target = "int_rate")
str(mlr3task) # -> R6 class object
mlr3task

# Modify task to only contain our predictor variables of interest
# -> "Mutate" the Task
# Object$select -> this is a method  !!!
mlr3task$select(c("loan_amnt", "annual_inc", "numeric_grade", "home_ownership",
                "purpose"))
mlr3task

library("mlr3viz")
autoplot(mlr3task)
autoplot(mlr3task, type = "pairs")
```

## Instantiate learner
```{r}
# Look up all available learners
library("mlr3learners")
available_learners <- as.data.table(mlr_learners)
available_learners
available_learners[task_type == "regr", c("key", "packages")]
rm(available_learners)

# We will use the learner "regr.lm"
# Instantiation:
mlr3learner = mlr_learners$get("regr.lm") # short cut: mylearner = lrn("regr.lm")
str(mlr3learner) # -> R6 class object
mlr3learner

# check pre-defined parameters for this learner
mlr3learner$param_set
```

## Train the learner
```{r}
# Define train and test rows
set.seed(123)
mlr3splits = partition(mlr3task, ratio = 0.8)
str(mlr3splits)

# Object$select -> this is a method  !!!
mlr3learner
mlr3learner$train(mlr3task, mlr3splits$train)
mlr3learner

mlr3learner$model

# for comparison
lm.fit
```

## Evaluation
```{r}
# Look for available performance measures
mlr_measures
mlr_measures$keys("regr")

# we will again use R Squared and the Root Mean Squared Error (RMSE)
mlr3measures = mlr_measures$get("regr.rsq")
mlr3measures = msr("regr.rsq") # short form !
mlr3measures = msrs(c("regr.rsq", "regr.rmse")) # short form with > 1 measure
mlr3measures

# Make prediction
mlr3predict = mlr3learner$predict(mlr3task, mlr3splits$test)
mlr3predict

# Compute selected performance measures
mlr3predict$score(mlr3measures)

# plot true versus true
autoplot(mlr3predict)
```
> regr.rsq regr.rmse 
>0.9024292 1.1644643



********************************************************************************
# tidymodels
********************************************************************************

## load packages
```{r}
library(tidyverse)
library(tidymodels)
```


## Load data (again)
```{r}
data_url <- "https://raw.githubusercontent.com/staehlo/mlr3_and_tidymodels/main/example_data_set.csv"
loans2 <- read_csv(file = data_url)
loans2 <- loans2 %>%
  mutate(numeric_grade = match(grade, LETTERS),
         home_ownership = factor(home_ownership),
         purpose = factor(purpose)
         ) %>%
  select(-c(grade, term)) %>%
  filter(annual_inc < 200000 & home_ownership != "NONE")
loans2
```

## Data Splitting
```{r}
set.seed(123)
tidysplit <- initial_split(loans2, prop = 0.8)
str(tidysplit)

tidytrain <- training(tidysplit)
tidytest <- testing(tidysplit)
tidytrain
```

## Create recipe
```{r}
tidyrecipe <- recipe(int_rate ~ loan_amnt + annual_inc + numeric_grade +
                       home_ownership + purpose, data = tidytrain)
tidyrecipe
summary(tidyrecipe)
```

## Create learner
```{r}
# Look up available learners: https://www.tidymodels.org/find/parsnip/#models
linear_reg()
tidylearner <- linear_reg() %>% set_engine("lm") # set_engine here not necessary
str(tidylearner)
```

## Build model
```{r}
tidyworkflow <- workflow() %>% add_model(tidylearner) %>% add_recipe(tidyrecipe)
tidyfit <- tidyworkflow %>% fit(data = tidytrain)

str(tidyfit)
tidyfit
tidyfit %>% extract_fit_parsnip() %>% tidy
```

## Evaluation
```{r}
# we will again use R Squared and the Root Mean Squared Error (RMSE)
tidymetrics <- metric_set(rsq, rmse)
tidymetrics

# Make prediction
tidypredict <- predict(tidyfit, new_data = tidytest)
tidypredict

# Compute selected performance measures
tidymetrics(data = tidytest, truth = int_rate, estimate = tidypredict$.pred)

ggplot(mapping = aes(x = tidypredict$.pred, y = tidytest$int_rate)) +
  geom_point() + xlab("predicted interest rate") + ylab("true interest rate") +
  geom_abline(intercept = 0, slope = 1, color = "blue")
```
> # A tibble: 2 × 3
>   .metric .estimator .estimate
>   <chr>   <chr>          <dbl>
> 1 rsq     standard       0.902
> 2 rmse    standard       1.16 
