---
title: "04 Classification"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# some file setup parameters
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999) # display numbers in non-scientific format
```

(2) Classification: Term has only two values: 36 and 60 month (3 or 5 years)
                    Train a model to predict Lending Club's proposed payback times.
              (based on loan amount, home ownership, annual income, and purpose)

Dependent variable:

* term (either 3 or 5 years)

Predictor variables:

* loan_amnt (integer)
* annual_inc (numeric)
* numeric_grade (integer)
* home_ownership (factor)
* purpose (factor)
* int_rate (numeric)

## Load data (again)
```{r}
library("tidyverse")
data_url <- "https://raw.githubusercontent.com/staehlo/mlr3_and_tidymodels/main/example_data_set.csv"
loans <- read_csv(file = data_url) %>%
  mutate(numeric_grade = match(grade, LETTERS),
         home_ownership = factor(home_ownership),
         purpose = factor(purpose),
         term = factor(term)
         ) %>%
  select(-c(grade, int_rate)) %>%
  filter(annual_inc < 200000 & home_ownership != "NONE")
rm(data_url)
loans
```


********************************************************************************
# Standalone: Using the randomForest package
********************************************************************************

```{r}
library(randomForest)

# Define test and training data sets
set.seed(123)
rows_for_training <- sample(nrow(loans), 0.8 * nrow(loans))
train <- loans[rows_for_training,]
test <- loans[-rows_for_training,]

# mtry
# ------------------------------------------------
# mtry = number of predictors used at each node
# rule of thumb: mtry = sqrt(number of predictors)
# mtry = example for a hyperparameter !

# importance
# ------------------------------------------------
# The permutation feature importance method would be used to determine the effects of the
# variables in the random forest model. This method calculates the increase in the
# prediction error(MSE) after permuting the feature values. If the permuting wouldn’t
# change the model error, the related feature is considered unimportant.
# (https://www.r-bloggers.com/2021/07/feature-importance-in-random-forest/)

start <- Sys.time()
rf.fit <- randomForest(term ~ loan_amnt + annual_inc + numeric_grade +
                         home_ownership + purpose, ntree = 100,
                       data = train, mtry = sqrt(6), importance = TRUE)
end <- Sys.time()
end - start; rm(start, end)

rf.fit
plot(rf.fit)
varImpPlot(rf.fit)

rf.predict <- predict(rf.fit, newdata = test)
head(rf.predict)

# confusion matrix
table(rf.predict, test$term)

# accuracy rate
1 - mean(rf.predict != test$term)

# error rate
mean(rf.predict != test$term)

rm(rows_for_training, test, train)
```
> end - start; rm(start, end)
Time difference of 5.9082 secs
> # accuracy rate
[1] 0.8024312
> # error rate
[1] 0.1975688


********************************************************************************
# mlr3
********************************************************************************

## Load package
```{r}
# install.packages("mlr3verse")
library("mlr3")
library("mlr3viz")
library("mlr3learners")
```

## Construct task
```{r}
mlr3task = as_task_classif(loans, target = "term")
# mlr3task$select(...) -> not necessary as we want to use all columns
autoplot(mlr3task)
autoplot(mlr3task, type = "pairs")
```

## Instantiate learner
```{r}
available_learners <- as.data.table(mlr_learners)
available_learners[task_type == "classif", c("key", "packages")]
rm(available_learners)

# We will use the learner "classif.ranger"
library(ranger) # Instanitation of learner requires loaded package !
# Option 1:
mlr3learner = mlr_learners$get("classif.ranger") # 
# Option 2: better, as you can directly set the hyperparameters
mlr3learner = lrn("classif.ranger", mtry = round(sqrt(6)), num.trees = 100)
mlr3learner
```

## Train the learner
```{r}
# Define train and test rows
set.seed(123)
mlr3splits = partition(mlr3task, ratio = 0.8)

start = Sys.time()
mlr3learner$train(mlr3task, mlr3splits$train)
end = Sys.time()
end - start; rm(start, end)

mlr3learner
mlr3learner$model
```

## Evaluation
```{r}
# Look for available performance measures
mlr_measures$keys("classif")

# We will use accuracy ("acc") and classification error ("ce")
mlr3measures = msrs(c("classif.acc", "classif.ce")) # short form for > 1 measure

# Make prediction
mlr3predict = mlr3learner$predict(mlr3task, mlr3splits$test)

# Compute selected performance measures
mlr3predict$score(mlr3measures)

# plot true versus true
autoplot(mlr3predict)
```
> end - start; rm(start, end)
Time difference of 3.075533 secs
> mlr3predict$score(mlr3measures)
classif.acc  classif.ce 
  0.7980548   0.2019452


********************************************************************************
# tidymodels
********************************************************************************

## load packages
```{r}
library(tidymodels)
```

## Data Splitting
```{r}
set.seed(123)
tidysplit <- initial_split(loans, prop = 0.8)
tidytrain <- training(tidysplit)
tidytest <- testing(tidysplit)
```

## Create recipe
```{r}
tidyrecipe <- recipe(term  ~ loan_amnt + annual_inc + numeric_grade +
                       home_ownership + purpose, data = tidytrain)
```

## Create learner
```{r}
# Look up available learners: https://www.tidymodels.org/find/parsnip/#models
tidylearner <- rand_forest(mtry = sqrt(6), trees = 100) %>%
  set_mode("classification") %>% # this wasn't necessary for linear_reg()
  set_engine("ranger")
```

## Build model
```{r}
tidyworkflow <- workflow() %>% add_model(tidylearner) %>% add_recipe(tidyrecipe)

start = Sys.time()
tidyfit <- tidyworkflow %>% fit(data = tidytrain)
end = Sys.time()
end - start; rm(start, end)

tidyfit
```

## Evaluation
```{r}
tidymetrics <- metric_set(accuracy)

# Make prediction
tidypredict <- predict(tidyfit, new_data = tidytest)

# Compute selected performance measures
tidyevaluation <- tidymetrics(data = tidytest, truth = term,
                              estimate = tidypredict$.pred_class)

# Accuracy
tidyevaluation$.estimate

# Error
1 - tidyevaluation$.estimate
```
> end - start; rm(start, end)
Time difference of 3.130244 secs
> tidyevaluation$.estimate (# Accuracy)
[1] 0.8028151
> 1 - tidyevaluation$.estimate (# Error)
[1] 0.1971849
